Spatial Analysis

### 2021-06-01 Scripts/spatial_analisis.Rmd
Looks like... its... done?

---
title: "Spatial Analysis"
output: html_notebook
---

```{r setup, include=FALSE}
MyFunctions::my_lib(c("ggmap","sf","tidyverse","tools","readr","data.table","maps"))
```

@@

# us_state_w <- rgdal::readOGR(dsn = "~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb")
state_sf <-  st_read("~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb") %>% 
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  rename(state = jurisdicti)
  # st_transform(crs = 4326)  # for matching projections
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

@@

@@

@@

# Plot to make sure makes sense (Picasso syle)
grid_eez_sf %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>% #0.1 for paper
  ggplot() +
  geom_sf(aes(color = state))

@@

### Map of smapling points

@@

```{r survey_map, eval = T, echo = T, fig.width = 12, fig.height = 8}

ggplot(us_map) +
  geom_sf() +
  geom_sf(data = World_eez_sf, aes(), fill = "white") +
  geom_point(data = subset(ocean_data, wtcpue = 0),
             aes(
               x = lon,
               y = lat
             ),
             color = "grey95",
             size = 1
  ) +
  geom_point(data = subset(ocean_data, wtcpue > 0),
             aes(
               x = lon,
               y = lat,
               color = log10(wtcpue)
             ),
             size = 1
  ) +
  scale_color_distiller(palette = "Spectral", 
                       guide_legend(title = "WCPUE per Haul (log10)")) + 
  coord_sf(xlim = c(-76, -65),ylim = c(35, 45)) +
  MyFunctions::my_ggtheme_m() +
  facet_wrap(~region)

@@

## SF example A

https://swilke-geoscience.net/post/spatial_interpolation/

```{r interpolation, eval = T, echo = T}

subset_data <- ocean_data %>% 
  filter(region == "Northeast US Spring",
         wtcpue > 0) %>% 
  group_by(lon,lat) %>% 
  summarise(wtcpue = sum(wtcpue,na.rm = T))

# First let's define a bounding Box to extrapolate from
bbox <- c(
  "xmin" = min(subset_data$lon),
  "ymin" = min(subset_data$lat),
  "xmax" = max(subset_data$lon),
  "ymax" = max(subset_data$lat)
)

# Expand the grid
grd_template <- expand.grid(
  lon = seq(from = bbox["xmin"], to = bbox["xmax"], by = 0.5),
  lat = seq(from = bbox["ymin"], to = bbox["ymax"], by = 0.5)
)

# ---------------- #
# Visualize the box

# ggplot() +
#   geom_point(data = grd_template, aes(x = lon, y = lat), size = 0.1) +
#   geom_point(data = subset_data,
#   mapping = aes(x = lon, y = lat, color = wtcpue), size = 3) +
#   scale_color_gradientn(colors = c("blue", "yellow", "red")) +
#   theme_bw()

# ---------------- #

# Rasterize our grid
# expects a PROJ.4 string, see https://epsg.io/25833

crs_raster_format <- "+proj=longlat +datum=WGS84 +no_defs"

grd_template_raster <- grd_template %>% 
  dplyr::mutate(Z = 0) %>% 
  raster::rasterFromXYZ( 
    crs = crs_raster_format)

# The model

# Triangular Irregular Surface
fit_tin <- interp::interp( # using {interp}
  x = subset_data$lon,           # the function actually accepts coordinate vectors
  y = subset_data$lat,
  z = subset_data$wtcpue,
  xo = grd_template$lon,     # here we already define the target grid
  yo = grd_template$lat,
  output = "points"
) %>% 
  bind_cols() %>% 
  filter(!is.na(z))

ggplot(us_map) +
  geom_sf()+
  geom_sf(data = World_eez_sf, aes(), fill = "white")+
  geom_tile( data = fit_tin,
             aes(
               x = x,
               y = y,
               fill =z,
               colour = z
             )
  ) +
  geom_point(data = subset(subset_data, wtcpue > 0),
             aes(
               x = lon,
               y = lat
             ),
             alpha = 0.2
  ) +
  scale_color_distiller(palette = "Spectral", 
                        guide_legend(title = "WCPUE per Haul")) + 
  scale_fill_distiller(palette = "Spectral", 
                       guide_legend(title = "WCPUE per Haul")) +
  coord_sf(xlim = c(-76, -65),ylim = c(35, 45)) +
  MyFunctions::my_ggtheme_m()

@@

>>>>

---
title: "Spatial Analysis"
# output: html_notebook
output: html_document
---

```{r setup, include=FALSE}
MyFunctions::my_lib(c("ggmap","sf","tidyverse","tools","readr","data.table","maps","viridis"))
library(MyFunctions)
```

@@

# us_state_w <- rgdal::readOGR(dsn = "~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb")
state_sf <-  st_read("~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  rename(state = jurisdicti) %>% 
  # st_transform(crs = 4326)  # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

grid_sf <-  st_read("/Volumes/Enterprise/Data/AcrossBoundaries/Data/Spatial/grid_sf/grid_eez_sf.shp")

@@

## Create basis sf

In this first step we first combine the North East EEZ shapefile (SAU) with the state waters sf provided by [NOAA](link) and girded the area following these steps:

- 1. Expand state waters using a buffer

- 2. Grid that buffer on a 0.3 ressolution

- 3. Crop the buffer to the EEZ

@@

# write_sf(grid_eez_sf, "../Results/Partial/grid_eez_sf.shp")

@@

# write_csv(grid_eez_df, "../Results/Partial/grid_eez_df.csv")

# Plot to make sure makes sense (Picasso syle)
# grid_eez_sf %>%
#   st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>% #0.1 for paper
#   ggplot() +
#   geom_sf(aes(color = state))

@@

## Map of smapling points

@@

```{r survey_map, eval = F, echo = T, fig.width = 12, fig.height = 8}

ggplot(land_sf) +
  geom_sf() +
  geom_sf(data = eez_sf, aes(), fill = "white") +
  geom_point(data = subset(ocean_data, wtcpue = 0),
             aes(
               x = lon,
               y = lat
             ),
             color = "grey95",
             size = 1
  ) +
  geom_point(data = subset(ocean_data, wtcpue > 0),
             aes(
               x = lon,
               y = lat,
               color = log10(wtcpue)
             ),
             size = 1
  ) +
  scale_color_distiller(palette = "Spectral", 
                       guide_legend(title = "WCPUE per Haul (log10)")) + 
  coord_sf(xlim = c(-76, -65),ylim = c(35, 45)) +
  MyFunctions::my_ggtheme_m() +
  facet_wrap(~region)

@@

Here we [interpolate](https://swilke-geoscience.net/post/spatial_interpolation/) the survey data within the grid. We

## Prepare data

- We removed cases where `wtcpue < 0`

```{r prepare_data, eval = T, echo = T}

# Subset data
subset_data <- ocean_data %>% 
  filter(
    region == "Northeast US Spring",
    wtcpue > 0)

# Get grid
grd_template <- data.frame(index = grid_sf$index,
                       state = grid_sf$state,
                       st_coordinates(grid_sf))

@@

## Fun. interpol

This is the main function used to interpolate the data per year. It follows a Triangular Irregular Surface method using the `interp::interp()` function.

```{r interpol_function, eval = T, echo = T}

tis <- function(input_data,yr){

  # print(paste(yr))

  data <- input_data %>% 
    filter(year == yr)

  if(nrow(data) == 0){
    return(tibble())
  }else{
    # Triangular Irregular Surface
    fit_tin <- interp::interp( # using {interp}
      x = data$lon,           # the function actually accepts coordinate vectors
      y = data$lat,
      z = data$wtcpue,
      xo = grd_template$X,     # here we already define the target grid
      yo = grd_template$Y,
      output = "points"
    ) %>% 
      bind_cols() %>% 
      bind_cols(grd_template) %>%
      mutate(year = yr) %>% 
      select(index, state, year, lon=x, lat=y, value = z)

    return(fit_tin)
  }
}

## Control pannel

```{r contro_pannel, eval = T, echo = T}

# Run interpolation for all years
years <- seq(1973,2019,1)

historic_tif <- bind_rows(
  lapply(years,tis,input_data = subset_data)
)

write_csv(historic_tif,
          paste0(my_path("R",),"interpolated_data.csv"))

# Results

```{r load_data}

# Periods
periods <-tibble(
  period = c(rep("a early",12),
             rep("b mid",12),
             rep("c late",12),
             rep("d now",12)
             ),
  year = c(seq(1972,1984,1),
            seq(1985,1997,1),
            seq(1998,2011,1),
            seq(2012,2019,1)
            )
  )

## Points

```{r area_plot, eval = T, echo = T}

 data_grid <- historic_tif %>% 
  # left_join(periods) %>% 
  # group_by(state,lat,lon,period) %>% 
  group_by(state,lat,lon) %>% 
  summarise(mean = mean(value,na.rm= T), .groups = "drop") %>% 
  rename(ID = state)

ggplot(data_grid) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  facet_wrap(~ ID) +
  scale_fill_viridis("Mean Proportion", alpha = 0.8) +
  scale_color_viridis("Mean Proportion", alpha = 0.8)

## Map

```{r area_plot, eval = T, echo = T}

total_fited <- historic_tif %>% 
  group_by(year) %>% 
  summarise(total_value = sum(value,na.rm=T))

state_fit <- historic_tif %>% 
  group_by(state,year) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = "year") %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods) %>% 
  group_by(ID = state,period) %>% 
  summarise(mean_per = round(mean(percentage)))

# check percentages
# state_fit %>% 
#   group_by(period) %>% 
#   summarise(sum(mean_per))

land_sf %>% 
  left_join(state_fit,
            by = "ID") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Mean Proportion", alpha = 0.8) +
  facet_wrap(~period, nrow = 1) +
  my_ggtheme_m()

## Area plot

This graph shows the proportion of each State over time

```{r area_plot, eval = T, echo = T}

total_fited <- historic_tif %>% 
  group_by(year) %>% 
  summarise(total_value = sum(value,na.rm=T))

# group by state
state_fit <- historic_tif %>% 
  group_by(state,year) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = "year") %>%
  mutate(percentage = state_value/total_value*100)

p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(percentage),
      fill = state
    )
  ) +
  viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  MyFunctions::my_ggtheme_p()

plotly::ggplotly(p)
